---
title: "HW1"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}

library(tidyverse) #read csv
```

## Problem 1

```{r include = FALSE}

data("penguins", package = "palmerpenguins")
penguins_df = penguins
```

***Penguin Dataset Description***

The penguins dataset is described by **`r ncol(penguins_df)`** variables, which are: **`r colnames(penguins_df)`**. This dataset is **`r nrow(penguins_df)`** rows by **`r ncol(penguins_df)`** columns. The average flipper length of the penguins in this dataset is **`r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`** mm.

***Penguin Dataset Scatter Plot***

Below is the scatter plot of the penguin flipper length against the penguin bill length. They are sorted by species and lengths are in mm:

```{r echo = FALSE, message = FALSE, warning = FALSE}
plot_penguins_df = tibble (
  x = pull(penguins, bill_length_mm),
  y = pull(penguins, flipper_length_mm)
)
ggplot (plot_penguins_df, aes(x = pull(penguins, bill_length_mm), y = pull(penguins, flipper_length_mm), col=penguins_df$species)) + geom_point() + labs(x = "Bill Length (mm)", y = "Flipper Length (mm)", title = "Penguin Flipper Length vs Bill Length (in mm)", color = "Penguin Species") +  theme(plot.title = element_text(hjust = 0.5))
  
```